---
title: "p8105_hw6_mk5149"
output: github_document
date: "2025-11-25"
author: "Mungyu Kwok"
---

# Preparation
Load any necessary libraries for this homework
```{r, include=FALSE}
library(tidyverse)
library(broom)
```

# Problem 1
## Problem 1 Part 1: Data loading and cleaning
The analysis begins by loading the raw data `homicide-data.csv`. The data is cleaned according to the problem specifications: creating a `city_state` variable, creating a binary solved variable indicating whether the homicide is solved, omitting specified cities (Dallas, TX; Phoenix, AZ; and Kansas City, MO; Tulsa, AL), limiting to `White` and `Black` victims, and ensuring `victim_age` is numeric.

```{r}
# Load the data
homicides = read_csv("data/homicide-data.csv") 

homicides_cleaned = homicides |>
  # Step 1: create city_state variable
  mutate(city_state = paste(city, state, sep = ", "),
         # Step 2: create a binary solved variable: 1 for closed by arrest, 0 for unresoved
         solved = if_else(disposition == "Closed by arrest", 1, 0),
         # Step 3: ensure victim_age is numerice
         victim_age = as.numeric(victim_age)
  ) |>
  # Step 4: omit specified cities
  filter(!city_state %in% c("Dallas, TX", "Phoenix, AZ", "Kansas City, MO", "Tulsa, AL")) |>
  # Step 5: limit to white or black victims only
  filter(victim_race %in% c("White", "Black")) |>
  # Step 6: remove rows with missing age
  drop_na(victim_age) |>
  # Step 7: Convert victim_sex to a factor, setting 'Female' as the reference level
  mutate(victim_sex = fct_relevel(victim_sex, "Female"))

# Check "solved" binary variable
homicides_cleaned |> 
  pull(solved) |> 
  unique()

homicides_cleaned |>
  head()
  
```

## Problem 1 Part 2: Logiestic regression for Baltimore, MD
For the city of Baltimore, MD, use the `glm` function to fit a logistic regression with resolved vs unresolved as the outcome and victim age, sex and race as predictors. 
Save the output of `glm` as an R object; apply the `broom::tidy` to this object; and obtain the estimate and confidence interval of the adjusted odds ratio for solving homicides comparing male victims to female victims keeping all other variables fixed.
```{r}
# Filter data fro Baltimore, MD 
baltimore_df = homicides_cleaned |>
  filter(city_state == "Baltimore, MD")

# Fit the logistic regression model
glm_baltimore = glm(solved ~ victim_age + victim_sex + victim_race, 
                    data = baltimore_df,
                    family = binomial)

# Tidy the results
# Exponentiate to get ORs, extracts the required values
or_baltimore = tidy(glm_baltimore, conf.int = TRUE, exponentiate = TRUE) |>
  filter(term == "victim_sexMale") |>
  select(term, estimate, conf.low, conf.high)

or_baltimore
```
The adjusted odds ratio for solving homicides comparing male victims to female victims in Baltimore, MD is approximately $\mathbf{0.426}$, with a $95\%$ confidence interval of $(\mathbf{0.325}, \mathbf{0.558})$. This indicates that the odds of a homicide being solved for male victims are significantly lower than for female victims, controlling for age and race.

```{r}


```







